---
title: "Artist's Data Exploration & Visualization"
output: github_document
author: "Thao Tran"
date: "2024-01-15"
editor_options: 
  markdown: 
    wrap: 72
    
---

Project Description:
Acting as a well-known artist’s/band’s manager, using social media analytics to help improve their popularity.

---

Set up:

```{r setup, include=FALSE}
# set options for R markdown knitting
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(linewidth=80)
```

```{r knitr line wrap setup, include=FALSE}
# set up line wrapping in MD knit output
library(knitr)
hook_output = knit_hooks$get("output")
knit_hooks$set(output = function(x, options) 
{
 # this hook is used only when the linewidth option is not NULL
 if (!is.null(n <- options$linewidth)) 
 {
 x = knitr:::split_lines(x)
 # any lines wider than n should be wrapped
 if (any(nchar(x) > n)) 
 x = strwrap(x, width = n)
 x = paste(x, collapse = "\n")
 }
 hook_output(x, options)
})
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## INSTALL NECESSARY PACKAGES

```{r}

install.packages("remotes")
library(remotes)

# install GitHub version of vosonSML 0.32.10
install_github("vosonlab/vosonSML")
library(vosonSML)

# install GitHub version of rtweet 1.1.0.9001
install_github("ropensci/rtweet")

# Load other packages required for this session into library
library(magrittr)
library(igraph)
library(tidyr)
library(tidytext)
library(stopwords)
library(dplyr)

# get twitter App name, API key, API secret, Access token, Access token secret
my_app_name <- "Class7230ICT"
my_api_key <- "5k06tQOIdyTsT9xR914kY3epo"
my_api_secret <- "FsLe2v57PVk1L8Pt9ez95KQ0ROPMrOJh5SySyJRlwOwDfMfRum"
my_access_token <- "1632183185256878080-DgxyxOrMCJcKZ38xXb1gQSs8svlQBo"
my_access_token_secret <- "euv2yxyqjLlDP8CNWttAPGqWJ07nOArFNPRmZsLJlOaPh"

```

#-----------------------------------------------------------------------------#

## COLLECT DATA

To comprehensively analyze Justin Bieber's fan engagement and potential strategic
improvements, I strategically selected keywords reflecting fan behaviors and
artist activities. Leveraging Twitter API access, I collected the most recent
1000 tweets and retweets containing search terms such as "Justin Bieber" and
"Justin Bieber tour," focusing on fans' interest in his live performances and
global tours.

```{r}
# Search terms topic 1 - FAN ----
twitter_fan_data <- Authenticate("twitter",
                             appName = my_app_name,
                             apiKey = my_api_key,
                             apiSecret = my_api_secret,
                             accessToken = my_access_token,
                             accessTokenSecret = my_access_token_secret) %>%
  Collect(searchTerm = "Justin Bieber OR Justin Bieber tour",
          searchType = "recent",
          numTweets = 1000,
          lang = "en",
          includeRetweets = TRUE,
          writeToFile = TRUE,
          verbose = TRUE)
head(twitter_fan_data$tweets)
```

Expanding the dataset, I incorporated keywords like "Justin Bieber album" and
"Justin Bieber hits" to gauge audience reactions to his music releases, aiming
to identify popular albums and understand the factors contributing to their success.

```{r}
# Search terms topic 2 - MUSIC PRODUCT ----
twitter_musicProduct_data <- Authenticate("twitter",
                                 appName = my_app_name,
                                 apiKey = my_api_key,
                                 apiSecret = my_api_secret,
                                 accessToken = my_access_token,
                                 accessTokenSecret = my_access_token_secret) %>%
  Collect(searchTerm = "Justin Bieber album OR Justin Bieber hit",
          searchType = "recent",
          numTweets = 1000,
          lang = "en",
          includeRetweets = TRUE,
          writeToFile = TRUE,
          verbose = TRUE)
head(twitter_musicProduct_data$tweets)
```

Additionally, I explored collaboration dynamics with keywords "Justin Bieber
collab" and "Justin Bieber ft," targeting fans of artists previously collaborated
with Bieber. The outcomes aim to assess the success of these collaborations and
provide insights for potential future partnerships, offering valuable strategies
for enhancing Justin Bieber's overall popularity.

```{r}
# Search terms topic 3 - INTERESTED USERS ----
twitter_relatedFan_data <- Authenticate("twitter",
                                 appName = my_app_name,
                                 apiKey = my_api_key,
                                 apiSecret = my_api_secret,
                                 accessToken = my_access_token,
                                 accessTokenSecret = my_access_token_secret) %>%
  Collect(searchTerm = "Justin Bieber collab OR Justin Bieber ft",
          searchType = "recent",
          numTweets = 1000,
          lang = "en",
          includeRetweets = TRUE,
          writeToFile = TRUE,
          verbose = TRUE)
head(twitter_relatedFan_data$tweets)
```

#-----------------------------------------------------------------------------#

## ACTOR NETWORK

In order to identify the top 5 most influential users for the artist, I made an
actor network for each topic of search terms I created. The data collection process
provided data on important fans of Justin Bieber in the first actor network; revealed
top users who care about the artist’s music products in the second network; and last
but not least the third actor network pointed out the top 5 most influential interested
users who may not be a fan but know about Justin Bieber and support his collaborations
with other artists.

From each search topic, we can find different related users. I pick out the top 2
accounts from each list to perform further analysis as it can reveal various new
aspects the artist may want to pay attention to so he can reach more audiences
and increase his popularity in the future.

### Actor Network 1 - FAN

```{r}
# Create actor network and graph from the data
fan_network <- twitter_fan_data %>% Create("actor")
fan_graph <- fan_network %>% Graph()

# Write graph to file
write.graph(fan_graph, file = "FanNetwork.graphml", format = "graphml")

# Overwrite the 'name' attribute in the graph with the 'screen name' attribute
V(fan_graph)$name <- V(fan_graph)$screen_name

# Run Page Rank algorithm to find important users
rank_fan <- sort(page_rank(fan_graph)$vector, decreasing = TRUE)
head(rank_fan, n=5)
```
### Actor Network 2 - MUSIC PRODUCT

```{r}
musicProduct_network <- twitter_musicProduct_data %>% Create("actor")
musicProduct_graph <- musicProduct_network %>% Graph()

write.graph(musicProduct_graph, file = "MusicProduct.graphml", format = "graphml")

V(musicProduct_graph)$name <- V(musicProduct_graph)$screen_name
rank_musicProduct <- sort(page_rank(musicProduct_graph)$vector, decreasing = TRUE)
head(rank_musicProduct, n=5)
```
### Actor Network 3 - INTERESTED USERS

```{r}
relatedFan_network <- twitter_relatedFan_data %>% Create("actor")
relatedFan_graph <- relatedFan_network %>% Graph()

write.graph(relatedFan_graph, file = "RelatedFan.graphml", format = "graphml")

V(relatedFan_graph)$name <- V(relatedFan_graph)$screen_name
rank_relatedFan <- sort(page_rank(relatedFan_graph)$vector, decreasing = TRUE)
head(rank_relatedFan, n=5)
```
### CONCLUSION

Due to the lack of methods to collect further information, I was unable to find
many aspects these top users have in common other than the artist of choice. 
However, it is interesting to see what other interests they have. The information
we can get from the group of users related to search terms “Justin Bieber collab”
& “Justin Bieber ft“ is actually quite notable. The artist has not only influenced
his fan base but also reached some loyal fans of other artists from different
cultures. As mentioned earlier, over the past few years Justin Bieber has been
collaborating with many global artists who are outside of the U.S. music industry.
However, he has not featured any Korean artists yet. As I researched more on this
information, I noticed that there was a rumor of a collaboration between Justin
& a famous K-pop band, but no music product has been created afterward. It may be
the reason why K-pop fans started to show interest in our artist. That is an
example Justin Bieber can consider and if we perform further analysis, there may
be even more suggestions for more potential projects.

#-----------------------------------------------------------------------------#

## SEMANTIC NETWORK

### Semantic Network 1 - FAN
```{r}
semantic_fan_network <- twitter_fan_data %>% Create("semantic")
semantic_fan_graph <- semantic_fan_network %>% Graph()

write.graph(semantic_fan_graph, file = "FanSemantic.graphml", format = "graphml")

rank_semantic_fan <- sort(page_rank(semantic_fan_graph)$vector, decreasing = TRUE)
head(rank_semantic_fan, n=10)
```
Most of the terms are mentions the artist’s latest product – album “Justice” and
his world tour with the same name. That explains the appearance of the terms justice,
world, live, and hashtag #justice. On the other hand, two hashtags #peace and #guns
seem to be quite irrelevant to the artist. Some people used these hashtags in their
tweets about Bieber just to get more viewers since they are both popular terms on
social media. To sum up, most of the terms that appear together with the two keywords
can lead audiences to the most updated news about the artist in general and about
his live performances in particular.

### Semantic Network 2 - MUSIC PRODUCT
```{r}
semantic_musicProduct_network <- twitter_musicProduct_data %>% Create("semantic")
semantic_musicProduct_graph <- semantic_musicProduct_network %>% Graph()

write.graph(semantic_musicProduct_graph, file = "ProductSemantic.graphml", format = "graphml")

rank_semantic_musicProduct <- sort(page_rank(semantic_musicProduct_graph)$vector, decreasing = TRUE)
head(rank_semantic_musicProduct, n=10)
```
The results can be divided into two groups, the first one may contain terms such
as #music, #tunes, hit, and #hits which only open up general discussions on Justin
Bieber’s songs. Another group contains #radio, #nowplaying, #onlineradio, radio
which are used mostly by Twitter accounts of some online radio to attract Justin’s
audiences to their platforms. In general, terms that appear together with the 2
keywords only bring insight into some listeners’ reactions to the artist’s performance.

### Semantic Network 3 - INTERESTED USERS
```{r}
semantic_relatedFan_network <- twitter_relatedFan_data %>% Create("semantic")
semantic_relatedFan_graph <- semantic_relatedFan_network %>% Graph()

write.graph(semantic_relatedFan_graph, file = "RelatedSemantic.graphml", format = "graphml")

rank_semantic_relatedFan <- sort(page_rank(semantic_relatedFan_graph)$vector, decreasing = TRUE)
head(rank_semantic_relatedFan, n=10)
```
The prominent hashtag #vh1playlist represents audience requests for VH1 Playlist,
an Indian music television channel. By searching "justin bieber #Vh1Playlist,"
I've identified a promising market for Justin Bieber's music products through
daily fan requests, with crucial terms reflecting song names and strong engagement.
Noteworthy collaborations with artists like Dan & Shay, Ludacris, Chance the Rapper,
and Nicki Minaj continue to attract attention, offering potential inspiration for
future compositions. In contrast, hashtags #thegudtymezshow and #thejaythegreatshow,
specific to an online radio platform with minimal user engagement, provide limited
insights. The term "essence" refers to a song featuring Justin Bieber by Afropop
artists WizKid and Tems.

#-----------------------------------------------------------------------------#

## CALCULATE UNIQUE USERS

```{r}
# DATASET 1 - FAN

# Create a dataframe for the data set
dataFrame_twitter_fan <- data.frame(twitter_fan_data)
head(dataFrame_twitter_fan)

# Calculate unique users in the new dataframe just created using n_distinct() function from dplyr package
n_distinct(dataFrame_twitter_fan$tweets.user_id)
```

```{r}
# DATASET 2 - PRODUCT

# Create a dataframe for the data set
dataFrame_twitter_musicProduct <- data.frame(twitter_musicProduct_data)
head(dataFrame_twitter_musicProduct)

# Calculate unique users in the new dataframe just created using n_distinct() function from dplyr package
n_distinct(dataFrame_twitter_musicProduct$tweets.user_id)
```

```{r}
# DATASET 3 - INTERESTED USERS

# Create a dataframe for the data set
dataFrame_twitter_relatedFan <- data.frame(twitter_relatedFan_data)
head(dataFrame_twitter_relatedFan)

# Calculate unique users in the new dataframe just created using n_distinct() function from dplyr package
n_distinct(dataFrame_twitter_relatedFan$tweets.user_id)
```
### CONCLUSION

The results pointed out that even though I retrieved data from 1000 different tweets, nearly half of them
were from the same Twitter users.






